Molecular dynamics simulation of synchronization in driven particles

### 2021-01-08 manuscript/manuscript_ajp_2020.tex
this manuscript is getting really good!

\begin{abstract}
  %What problem did you study and why is it important?
  Particles
  %confined to a narrow channel 
  driven
  across a washboard potential energy landscape
  move in 
  synchronized patterns.   
  %exhibit a variety of synchronization effects.
  %in
  %  Numerical model
  We discuss a simple
  numerical model that
  %exhibits
  provides insight
  to synchronization behavior
  %and
  %atoms on surfaces.
  %voltage driven superconducting vortices
  %confined in a Josephson junction,
  %and ac and dc driven
  %charge and spin density waves.
  %We present the basics
  using
  molecular dynamics simulations
  for particles moving through
  a viscous liquid.
  This model reproduces behaviors observed
  in experimental systems 
  of %interest to condensed matter physics %including 
  magnetically driven colloidal particles
  confined by light-fields.
  %
  %applied to single and multiple particle systems.
  %so we model the dynamics of each particle
  %with 
  %overdamped equations of motion.
  Our results show a variety of synchronization effects 
  in single and multi-particle systems, 
  %and detect synchronization using
  %measurements of
  which we characterize with
  plots of 
  particle velocity as a function of applied driving force,
  and phase diagrams of position versus velocity.
  %We demonstrate how to 
  %visualize
  %the system with
  %numerical representations of the landscape
  %and
  %animations of the particle motion.
  %We include sample code
  %and exercises for students
  %with 
  %opportunities
  %to reproduce our results and propose
  %new numerical experiments.
  %With only a few particles in two-dimensions,
  %the simulation runs quickly,
  %making this an appropriate model for undergraduates to explore.  
\end{abstract}
% AJP requires an abstract for all regular article submissions.
% Abstracts are optional for submissions to the "Notes and Discussions" section.

@@

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we perform numerical studies 
on the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
Numerical studies %of harmonic systems
can yield insights
to the mechanics of naturally oscillating systems.
We chose this model for its
relevance to condensed matter systems
and ease of simulation.
%---------------------------------------------------------------
%kinks and nanofriction
%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In the following paper
In Section~\ref{sec:colloids} we describe colloid particles,
an relevent
experimental analog of our simulations.
We describe
our molecular dynamics model for a single particle in Section~\ref{sec:MD}.
%In this section, we restrict our discussion to a single particle,
%and include additional information
%for multiple particles in Section~\ref{sec:sync}.
%including various  and confining environments.
We summarize
our results 
%in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one}.
In Section~\ref{sec:sync}
we 
extend the model to include multiple particles,
and show how confined 
interacting particles respond to the same applied driving force.
%including stationary propagation of high density kinks
%in Sec.~\ref{sec:kink}.
%We include code to simulate
%and visualize the landscape and dynamics in this section
%and supplementary material.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
%We present these results using standard tools of non-linear oscillators
%such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWING COUPLING, OR QUASI-PERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
%In Section~\ref{sec:quasiperiod},
%we show how an aperiodic landscape modifies the particle dynamics.
%Finally we explore the transition to chaotic dynamics in 
%Section~\ref{sec:chaos}
%and conclude 
%in Section~\ref{sec:conclusion}. 
We conclude with
exercises for interested students 
in Section~\ref{sec:problems}.

@@

Complex dynamics can be studied with
colloid particles with experiments and simulation.
In experiment, colloids are typically spheres of 
plastic suspended in
highly de-ionized water or silica in suspended in organic solvent.
%experimental - colloids specifically
Because 
the spheres are a few microns in size
and move at rates of microns per second,
the position of a particle %is large and slow enough to
can be imaged as a function of time 
with a traditional optical camera.
%due to the large size and slow speed of the particles.
Individual colloids exhibit 
Brownian motion,
where particles change direction %in a seemingly random fashion
due to unseen 
collisions 
with particles comprising the suspending fluid \cite{}.
Brownian motion is a temperature and viscosity dependent
phenomena. \cite{}
[elaborate - move to associated problems?].

Due to the relative ease of
manipulation and imagining, 
colloids are used as experimental models
for systems relatively hard to access and visualize. 
%SAME SENTENCE TWICE
Collections of colloids can
be used to study the properties of solids and liquids
%comprised of smaller particles
%such as atoms \cite{} or electrons \cite{}
or more exotic systems 
such as cold atoms or electron gases \cite{Grier2003}.  
The interaction forces between colloids can be controlled 
by altering the chemistry of the suspending fluid
or surface ligands of the particles.
Thus colloids can 
be modified to mimic different
systems such as hard spheres interacting via contact forces \cite{},
dipole interactions \cite{},
and long range electrostatic interactions \cite{}.

@@

%because 
We model the 
particle dynamics 
with an overdamped
equation of motion.
%integrated with the Verlet method.
%The suspending fluid is highly viscous
%and exerts a damping force on the particles equivalent
%to a linear drag force.
%Overdamped dynamics is a common assumption in models of colloidal particles.
%This is an assumption that
%The term overdamped or critically damped indicates
%that an oscillator cannot undergo periodic
%undulations due to some external force. %such as
%suspension in a viscous fluid \cite{}.
In our colloidal model we assume
the particle is suspending in a viscous fluid
that provides friction, or damping.
In the overdamped model the
linear friction produced by this 
nonconservative force,
$\vec{F}_{drag} = -\eta \vec{v}_i$ is
sufficient so that
the acceleration of the particle is zero.
The friction coefficient $\eta$
is proportional to the fluid viscosity.
Such a model in which inertia is set to zero
is appropriate for small particles
moving through a high viscosity fluid
such as a colloid moving in water.
We discuss this further in 
Section~\ref{sec:problems} Exercise~\ref{ex:reynolds}. %for more information

Newton's second law for an individual particle
$m \vec{a}_i = \sum \vec{F}$ is simplified
by the assumption $\vec{a}_i$ is zero in overdamped model.
%Newton's second law can be rearranged %for the velocity
%to
The equation of motion for an isolated particle is
%solved for velocity 
\begin{equation}
  \eta \vec{v}_i = \vec{F}^l_{i} + \vec{F}^{D}(t).
    \label{eq:motion}
\end{equation}
with the viscosity coefficient $\eta = 1$,
which we describe in more detail in Table~\ref{tab:1}
and Exercise~\ref{ex:reynolds}.
In Section~\ref{sec:problems} Exercise~\ref{ex:n2l}
we invite the reader to confirm this result.

@@

\section{Synchronized particle chains}
\label{sec:sync}

@@

\item Brownian motion.  Consider the consequences of a non-zero temperature on the simulation, in which particles will collide with invisible particles making up the suspending fluid.  These collisions are more likely at higher temperature due to the increased kinetic energy of the fluid particles, as described by the Maxwell-Boltzmann distribution \cite{}.  In molecular dynamics simulations temperature effects can be modeled by including randomized forces $f^T$ to emulate the motion of particles undergoing Brownian motion.  A distribution of randomizes forces contributes equally in all directions such that the force averaged over a finite time interval is zero $\langle f^T(t) \rangle = 0$.  Random distributions of contain no correlations over the independent variable (time in this case) $\langle f^T(t) f^T(\tau)\rangle = 2 \eta k_B T \delta(t-\tau)$ where the energy $k_B T$ derived from the Boltzmann constant $k_B$ and temperature $T$ has energy units $E_0$.  \cite{Allen2017}

@@

\section{Conclusion}
\label{sec:conclusion}	% You can label sections for reference

@@

>>>>

\begin{abstract}
  Synchronization
  plays a key role in many physical processes.
  %Simplified models
  %such as 
  %What problem did you study and why is it important?
  Particles
  %confined to a narrow channel 
  driven
  across a washboard potential energy landscape
  move in 
  synchronized patterns,
  as observed in 
  experiments %al systems 
  of %interest to condensed matter physics %including 
  magnetically driven colloidal particles
  confined by light-fields.
  %exhibit a variety of synchronization effects.
  %in
  %  Numerical model
  We discuss a simple
  numerical model that
  %exhibits
  provides insight
  to synchronization behavior
  %and
  %atoms on surfaces.
  %voltage driven superconducting vortices
  %confined in a Josephson junction,
  %and ac and dc driven
  %charge and spin density waves.
  %We present the basics
  using
  molecular dynamics simulations
  for particles moving through
  a viscous liquid.
  %
  %applied to single and multiple particle systems.
  %so we model the dynamics of each particle
  %with 
  %overdamped equations of motion.
  Our results show a variety of synchronization effects 
  in single and multi-particle systems, 
  %and detect synchronization using
  %measurements of
  which we characterize with
  plots of 
  particle velocity as a function of applied driving force,
  and phase diagrams of position versus velocity.
  %We demonstrate how to 
  %visualize
  %the system with
  %numerical representations of the landscape
  %and
  %animations of the particle motion.
  %We include sample code
  %and exercises for students
  %with 
  %opportunities
  %to reproduce our results and propose
  %new numerical experiments.
  %With only a few particles in two-dimensions,
  %the simulation runs quickly,
  %making this an appropriate model for undergraduates to explore.  
\end{abstract}
% AJP requires an abstract for all regular article submissions.
% Abstracts are optional for submissions to the "Notes and Discussions" section.

@@

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we perform numerical studies 
on the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
%Numerical studies %of harmonic systems
%can yield insights
%to the mechanics of naturally oscillating systems.
We chose this model for its
relevance to condensed matter systems
and ease of simulation.
%---------------------------------------------------------------
%kinks and nanofriction
%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In the following paper
In Section~\ref{sec:colloids} we describe colloid particles,
an relevent
experimental analog of our simulations.
We describe
our molecular dynamics model for a single particle in Section~\ref{sec:MD}.
%In this section, we restrict our discussion to a single particle,
%and include additional information
%for multiple particles in Section~\ref{sec:sync}.
%including various  and confining environments.
We summarize
our results 
%in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one}.
In Section~\ref{sec:sync}
we 
extend the model to include multiple particles,
and show how confined 
interacting particles respond to the same applied driving force.
%including stationary propagation of high density kinks
%in Sec.~\ref{sec:kink}.
%We include code to simulate
%and visualize the landscape and dynamics in this section
%and supplementary material.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
%We present these results using standard tools of non-linear oscillators
%such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWING COUPLING, OR QUASI-PERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
%In Section~\ref{sec:quasiperiod},
%we show how an aperiodic landscape modifies the particle dynamics.
%Finally we explore the transition to chaotic dynamics in 
%Section~\ref{sec:chaos}
%and conclude 
In Section~\ref{sec:conclusion}
we describe how our work with colloids applies more broadly
to many physical systems
such as dusty plasmas, superconducting vortices and Josephson junctions.
We conclude with
exercises for interested students 
in Section~\ref{sec:problems}.

@@

Complex dynamics can be studied with
colloid particles with experiments and simulation.
In typical experiments colloids are spheres of 
plastic suspended in
highly de-ionized water or silica in suspended in organic solvent.
%experimental - colloids specifically
Because 
the spheres are a few microns in size
and move at rates of microns per second,
the position of a particle %is large and slow enough to
can be imaged as a function of time 
with a traditional optical camera.
%due to the large size and slow speed of the particles.
Individual colloids exhibit 
Brownian motion,
where particles change direction %in a seemingly random fashion
due to 
collisions 
with unseen particles comprising the suspending fluid \cite{}.
%[elaborate - move to associated problems?].

@@

%because
The inertia of 
small particles is greatly reduced by interactions
with fluid particles \cite{Purcell1977}.
Our model 
colloids are overdamped
-- i.e. suspended in a continuous viscous fluid
that dissipates energy at a rate such 
that the particles do not accelerate.
%that greatly impedes their motions.
Energy dissipation from the fluid is modeled
with a frictional force on particle with velocity $\vec{v}_i$
%that removes energy from
%the particles,
%which we model as
$\vec{F}^{drag}_i = -\eta \vec{v}_i$
%proportional to the particle velocity 
with a friction coefficient $\eta$
proportional to the fluid viscosity.
We discuss friction models for
spheres moving through fluids in 
Section~\ref{sec:problems} Exercise~\ref{ex:reynolds}. %for more information
%We model the particle dynamics  with an overdamped equation of motion. Models with inertia set to zero are appropriate for moving through a high viscosity fluid
Newton's second law for an individual particle
%$m \vec{a}_i = \sum \vec{F}$
is simplified
by the assumption $\vec{a}_i$ is zero. % in the overdamped model.
%Newton's second law can be rearranged %for the velocity
%to
The overdamped equation of motion for an isolated particle is
%solved for velocity 
\begin{equation}
  \eta \vec{v}_i = \vec{F}^l_{i} + \vec{F}^{D}(t).
    \label{eq:motion}
\end{equation}
with  $\eta = 1$.  
%which we describe in more detail in Table~\ref{tab:1}
%and Exercise~\ref{ex:reynolds}.
In %Section~\ref{sec:problems}
Exercise~\ref{ex:n2l}
we invite the reader to confirm this result.

@@

\section{Synchronized particle chains}
\label{sec:sync}
The interaction forces between colloids can be controlled 
by altering the chemistry of the suspending fluid
or surface ligands of the particles.
Thus colloids can 
be modified to mimic different
systems such as hard spheres interacting via contact forces \cite{},
dipole interactions \cite{},
and long range electrostatic interactions \cite{}.

@@

\item {\bf Brownian motion}
  is a temperature and viscosity dependent
  phenomena. \cite{}
  Consider the consequences of a non-zero temperature on the simulation, in which particles will collide with invisible particles making up the suspending fluid.  These collisions are more likely at higher temperature due to the increased kinetic energy of the fluid particles, as described by the Maxwell-Boltzmann distribution \cite{}.  In molecular dynamics simulations temperature effects can be modeled by including randomized forces $f^T$ to emulate the motion of particles undergoing Brownian motion.  A distribution of randomizes forces contributes equally in all directions such that the force averaged over a finite time interval is zero $\langle f^T(t) \rangle = 0$.  Random distributions of contain no correlations over the independent variable (time in this case) $\langle f^T(t) f^T(\tau)\rangle = 2 \eta k_B T \delta(t-\tau)$ where the energy $k_B T$ derived from the Boltzmann constant $k_B$ and temperature $T$ has energy units $E_0$.  \cite{Allen2017}

@@

\section{Conclusion}
\label{sec:conclusion}	% You can label sections for reference
Due to the relative ease of
manipulation and imagining, 
colloids are used as experimental models
for systems relatively hard to access and visualize. 
%SAME SENTENCE TWICE
Collections of colloids can
be used to study the properties of solids and liquids
%comprised of smaller particles
%such as atoms \cite{} or electrons \cite{}
or more exotic systems 
such as cold atoms or electron gases \cite{Grier2003}.

@@

    \bibitem{Purcell1977} E. M. Purcell, Life at low Reynolds numbers, Am. J. Phys. {\bf 45}, 3–11 (1977).
